---
title: "Practical Machine Learning Coursework"
author: "Weiguang Zhou"
date: "28 April 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

```{r cars, echo=TRUE, message=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
set.seed(428)
```

#Data Preprocessing
Firstly, combine the two CSV files into one. This guarantees that the data type of the training and texting data sets are the same.

Load the data
```{r}
#the following code is for the examiner
#training <- read.csv(url(trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), na.strings=c("NA","#DIV/0!",""))
#testing <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"), na.strings=c("NA","#DIV/0!",""))

#the following code is for me
#training <- read.csv("F:\\R_WorkDir\\Practical Machine Learning Course\\pml-training.csv", na.strings=c("NA","#DIV/0!",""))
#testing <- read.csv("F:\\R_WorkDir\\Practical Machine Learning Course\\pml-testing.csv", na.strings=c("NA","#DIV/0!",""))


#training=all[all$classe != 0,]
#testing=all[all$classe == 0,]
#index column is not needed.
#training$X =NULL
#testing$X = NULL

#this file is given in the github
all <- read.csv("F:\\R_WorkDir\\Practical Machine Learning Course\\pml-all.csv", na.strings=c("NA","#DIV/0!",""))
all$X = NULL
```

#Cleaning the data

```{r}
myDataNZV <- nearZeroVar(all, saveMetrics=TRUE)
myNZVvars <- names(all) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
all <- all[!myNZVvars]
dim(all)
```

```{r}
all.copy <- all #creating another subset to iterate in loop
for(i in 1:length(all)) { #for every column in the training dataset
        if( sum( is.na( all[, i] ) ) /nrow(all) >= .6 ) { #if n?? NAs > 60% of total observations
        for(j in 1:length(all)) {
            if( length( grep(names(all[i]), names(all.copy)[j]) ) ==1)  { #if the columns are the same:
                all.copy <- all.copy[ , -j] #Remove that column
            }   
        } 
    }
}
all = all.copy
```

#Split the data
```{r}
all.length=length(all$user_name)
training=head(all,all.length-20)
testing=tail(all,20)
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]; 
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```

#Train rpart
```{r rpart}
my.model <- rpart(classe ~ ., data=myTraining, method="class")
my.testing.result = predict(my.model, myTesting, type = "class")
confusionMatrix(my.testing.result, myTesting$classe)
```
#Train Random Forest
```{r random_forest}
my.rf <- randomForest(classe ~. , data=myTraining)
rf.predict <- predict(my.rf, myTesting, type = "class")
confusionMatrix(rf.predict, myTesting$classe)
```

#Predict
Compare the two models. Random forest is more accurate and it used.
```{r predict}
predictions <- predict(my.rf, testing, type = "class")
predictions
```
